---
title: "Customer Segmentation using RFM Analysis"
author: 'M Hendra Herviawan'

output:
  html_document:
    fig_height: 4
    fig_width: 7
---
## M Hendra Herviawan
## Project Overview
According to wikipedia.org, "RFM is a method used for analyzing customer value".

RFM stands for the three dimensions:

* Recency – How recently did the customer purchase?
* Frequency – How often do they purchase?
* Monetary Value – How much do they spend?

The resulting segments can be ordered from most valuable (highest recency, frequency, and value) to least valuable (lowest recency, frequency, and value). Identifying the most valuable RFM segments can capitalize on chance relationships in the data used for this analysis.
```{r, message=FALSE, warning=FALSE, results='hide'}
library(data.table)
library(dplyr)
library(ggplot2)
#library(stringr)
#library(DT)
library(tidyr)
library(knitr)
library(rmarkdown)
```

### Load Dataset
First, Lets we Load & Examine Dataset
```{r, results='markup', message=FALSE}
df_data <- fread('zcat /home/x/Documents/Kaggle/ecommerce-data/input/data.csv.zip')
glimpse(df_data)
```

### Data Cleaning
Delete all negative Quantity and Price. We also need to delete NA customer ID
```{r}
df_data <- df_data %>% 
  mutate(Quantity = replace(Quantity, Quantity<=0, NA),
         UnitPrice = replace(UnitPrice, UnitPrice<=0, NA))

df_data <- df_data %>%
  drop_na()

```
### Recode variables
We should do some recoding and convert character variables to factors.
```{r}
df_data <- df_data %>% 
  mutate(InvoiceNo=as.factor(InvoiceNo), StockCode=as.factor(StockCode), 
         InvoiceDate=as.Date(InvoiceDate, '%m/%d/%Y %H:%M'), CustomerID=as.factor(CustomerID), 
         Country=as.factor(Country))

df_data <- df_data %>% 
  mutate(total_dolar = Quantity*UnitPrice)

glimpse(df_data)
```

### Calculate RFM {.tabset}
To implement the RFM analysis, we need to further process the data set in by the following steps:

1. Find the most recent date for each ID and calculate the days to the now or some other date, to get the Recency data
2. Calculate the quantity of translations of a customer, to get the Frequency data
3. Sum the amount of money a customer spent and divide it by Frequency, to get the amount per transaction on average, that is the Monetary data.

```{r results='hold'}

df_RFM <- df_data %>% 
  group_by(CustomerID) %>% 
  summarise(recency=as.numeric(as.Date("2012-01-01")-max(InvoiceDate)),
            frequenci=n_distinct(InvoiceNo), monitery= sum(total_dolar)/n_distinct(InvoiceNo)) 

summary(df_RFM)

kable(head(df_RFM))
```

#### Recency
Recency – How recently did the customer purchase?
```{r} 
hist(df_RFM$recency)
```

#### Frequency
Frequency – How often do they purchase?
```{r} 
hist(df_RFM$frequenci, breaks = 50)
```

#### Monetary
Monetary Value – How much do they spend?
```{r} 
hist(df_RFM$monitery, breaks = 50)
```
Becouse the data is realy skewed, we use log scale to normalize
```{r}
df_RFM$monitery <- log(df_RFM$monitery)
hist(df_RFM$monitery)
```

### Clustering
```{r}
df_RFM2 <- df_RFM
row.names(df_RFM2) <- df_RFM2$CustomerID
df_RFM2$CustomerID <- NULL

df_RFM2 <- scale(df_RFM2)
summary(df_RFM2)
```

```{r}
d <- dist(df_RFM2)
c <- hclust(d, method = 'ward.D2')

plot(c)
```

#### Cut
```{r}
members <- cutree(c,k = 8)

members[1:5]
table(members)

aggregate(df_RFM[,2:4], by=list(members), mean)

```